Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-10-18 ms.tex
Edits to ms

There a number of caveats when considering the generality of our results.
First, the GPDD data represent a taxonomically and geographically biased sample
of populations. Also, the longer time series we focus on are dominated by
commercially and recreationally important species and a disproportionate number
of populations are located in the United Kingdom. Although we expect that we
would find strong but rare evidence for black swans in other large taxonomic or
geographic samples, the common forces driving those black swans (e.g.~severe
winters in Table~1) likely differ. Second, in a large database with disparate
data sources, we cannot ensure the quality of every data point. Recording
mistakes undoubtedly comprise some of the black swans we detected. Conversely,
recording mistakes could mask heavy tails if initial observations are discarded
or altered as assumed recording errors. Third, the temporal scale of observation
and population dynamics vary considerably across populations in the GPDD and
this likely influences the detection of heavy tails. For example, if we make
frequent observations relative to generation time (e.g.~for many large bodied
mammals) we will average across generations and perhaps not observe the tails.
Furthermore, if we record population censuses infrequently relative to
generation time (e.g.~many insects in the GPDD) the recorded data may average
across extreme and less-extreme events and dampen observed black swan dynamics.

Our results suggest a number of important future issues to address regarding
ecological black swans. First, given that black swans can and do occur, can we
forecast the probability of black swans in space and time? Furthermore, what
management policies allow us to detect them quickly after they happen?
Secondly, can we isolate the components of ecological dynamics that experience
black swans events by moving from phenomenological models such as the Gompertz
to mechanistic models, that take into account, for example, recruitment? We
expect that greater insight into the mechanisms and covariates of ecological
black swans may be best obtained through specific geographic and taxonomic
subsets of data where longer time series are available
\citep[e.g.][]{segura2013}. Finally, if a sizeable fraction of populations
experience heavy tails in population process noise, what is the impact of
allowing for black swans in forecasts of ecological risk? Financial analysts
are acutely concerned with the shape of the tail distribution of financial
returns since these impact estimates of risk --- the probability of a specific
magnitude of undesired event occurring (REF). It is possible that a comparable
focus in ecology would alter our perception of ecological forecasts such as
extinction risk.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{correlates-p10.pdf}
\caption{
Potential correlates of heavy-tailed behaviour (indicated by a high probability that $\nu < 10$).
Shown are (a--c) parameters from the Gompertz heavy-tailed population model, (b) number of time steps, (c) body length, and (d) lifespan.
For the Gompertz parameters, $\sigma$ refers to the scale parameter of the Student-t process-noise distribution, $\lambda$ refers to the population growth rate at an abundance of one, $b$ refers to the density dependence parameter ($1$ is maximally density independent, $0$ is maximally density dependent, and $<0$ is inversely density dependent).
%For parameter values, points and lines represent medians and 50\% credible intervals of the posterior.
Circles representing a limited number of sharks, crustaceans, and gastropods are filled in white.
Median and 90\% credible interval posterior predictions of a beta regression multilevel model are shown in panels a and d where there was a high probability the slope coefficient was different from zero (Supporting Material).
}
\label{fig:correlates}
\end{center}
\end{figure}

>>>>

There are a number of caveats when considering the generality of our results.
First, the GPDD data represent a taxonomically and geographically biased sample
of populations --- the longer time series we focus on are dominated by
commercially and recreationally important species and a disproportionate number
of populations are located in the United Kingdom. Although we expect that we
would find strong but rare evidence for black swans in other large taxonomic or
geographic samples, the common forces driving those black swans (e.g.~severe
winters in Table~1) likely differ. Second, in a large database with disparate
data sources, we cannot ensure the quality of every data point. Recording
mistakes undoubtedly comprise some of the black swans we detected. Conversely,
recording mistakes could mask heavy tails if initial observations assumed to be
recording errors and are discarded or altered. Third, the temporal scale of
observation and population dynamics vary considerably across populations in the
GPDD and this likely influences the detection of heavy tails. For example, if
we make frequent observations relative to generation time (e.g.~for many large
bodied mammals) we will average across generations and perhaps not observe
black swans. Furthermore, if we census populations infrequently relative to
generation time (e.g.~many insects in the GPDD) the recorded data may average
across extreme and less-extreme events and dampen black swan dynamics.

Our results suggest important future issues to address about ecological black
swans. Given that black swans can and do occur, can we forecast the probability
of black swans in space and time? Furthermore, what management policies allow
us to detect them quickly after they happen? Secondly, can we isolate the
components of ecological dynamics that experience black swans events by moving
from phenomenological models such as the Gompertz to mechanistic models, that
take into account, for example, recruitment? We expect that greater insight
into the mechanisms and covariates of ecological black swans may be best
obtained through specific geographic and taxonomic subsets of data where longer
time series are available \citep[e.g.][]{segura2013}. Finally, what is the
impact of allowing for black swans in forecasts of ecological risk? Financial
analysts are concerned with the shape of financial-return distribution tails
because these impact estimates of risk --- the probability of a specific
magnitude of undesired event occurring (REF). It is possible that a comparable
focus in ecology would alter our perception of ecological forecasts such as
extinction risk.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{correlates-p10.pdf}
\caption{
Potential correlates of heavy-tailed behaviour (indicated by a high probability
that $\nu < 10$). Shown are (a--c) parameters from the Gompertz heavy-tailed
population model, (b) number of time steps, (c) body length, and (d) lifespan.
For the Gompertz parameters, $\sigma$ refers to the scale parameter of the
Student-t process-noise distribution, $\lambda$ refers to the population growth
rate at an abundance of one, $b$ refers to the density dependence parameter
($1$ is maximally density independent, $0$ is maximally density dependent, and
$<0$ is inversely density dependent).
%For parameter values, points and lines represent medians and 50\% credible
%intervals of the posterior.
Circles representing a limited number of sharks, crustaceans, and gastropods
are filled in white. Median and 90\% credible interval posterior predictions of
a beta regression multilevel model are shown in panels a and d where there was
a high probability the slope coefficient was different from zero (Supporting
Material).
}
\label{fig:correlates}
\end{center}
\end{figure}
